Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-12-08 Cost_manuscript.Rmd
model discussion

  Our results provided mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/. Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that the normative predictions of traditional foraging models were insufficient to explain our results. Instead, we found that demands imposed a bias on the estimation of the opportunity cost of time, which converged over time if individuals were exposed to interleaving types of demands. We hypothesize that this bias arises from modulations of subjective time. For instance, cognitive effort could have become appealing by virtue of compressing the perceived trial time, a perception that could be recalibrated with exposure to alternative demands. The construal of such shorter experienced durations could be due to the ongoing recruitment of working memory (e.g. impeding ongoing estimations of the elapsed time), or because of the discrete nature of the cognitive task, which allowed individuals to count instead of thinking about time (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

@@

  Understanding the sporadic attractiveness of effort remains elusive. In this study, we tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Wikenheiser2013]. This successful integration of the task was further evidenced by the ability of individuals to describe near-optimal strategies after participation, discarding the possibility of our results being driven by noise or bias in the estimation of environmental richness [@Cash-Padgett2020; @Garrett2020; @Dundon2020]. In contrast, our prediction that effort requirements would produce more acceptances only held for the cognitive group, and this tendency for higher acceptances did not switch to passive waiting as predicted for the within-subjects experiment. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted task-specific attributes contributing to the attractive or aversive nature of demands. The fact that predicted stable preferences for cognitive effort carried over onto the beginning of the within-subjects experiment further supported a task-specific modulation. However, the unexpected gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discards the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that our population was mostly composed of demand seekers [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweis2018], as participants expressed their choices early and rarely quit during the handling time. Finally, the observed differences in acceptances between effort types cannot be explained by reward alone, as demands were rewarded at the same rate.

  (**hold off on this paragraph until modeling is done**) We do not discard the influence of these elements in daily effort evaluations. Instead, we computationally highlight that effort can became valuable by shortening the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands (Cognitive effort gained value by making trials feel shorter). Alterations in subjective time sensitivity have been shown to play a role in temporal discounting [@Mckerchar2009], and are often experienced when we are fully immersed in a task [@Csikszentmihalyi2014]. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios. ....This could be because: 1) discrete nature of cog taska llowed for counting; 2) flow; 3) persistent recruitment of working memory (Buhi and Gershman). Mention that order effects reflect the gradual integration of subjective time across demands in exp 2. Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort. Maybe use this example: if I ask you to perform an activity until you think a minute has elapsed, that estimate will be noisy, but also different for different tasks. That's how it is in real life.

@@

>>>>

  Our results provided mixed support for our pre-registered hypotheses, which are detailed in https://osf.io/2rsgm/. Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that the normative predictions of traditional foraging models were insufficient to explain our results. Instead, we found that demands imposed a bias on the estimation of the opportunity cost of time, which converged over time if individuals were exposed to interleaving types of demands. We hypothesize that this bias arises from modulations of subjective time. For instance, cognitive effort could have become appealing by virtue of compressing the perceived trial time, a perception that could be recalibrated with exposure to alternative demands. The construal of such shorter experienced durations could be due to the ongoing recruitment of working memory (e.g. preventing individuals from estimating the elapsed time), or because of the discrete nature of the cognitive task, which allowed individuals to count instead of thinking about time (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

@@

  Understanding the sporadic attractiveness of effort remains elusive. In this study we sought to identify factors underlying this phenomenon, testing the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced.

  Our findings provided partial support for our pre-registered hypotheses (https://osf.io/2rsgm/registrations). In accord with our predictions, decision makers in both experiments successfully approximated reward-maximizing strategies, as evidenced by acceptance rates near optimal points. This is in line with previous foraging experiments, where the main deviation from optimality has been a tendency to overharvest [@Constantino2015; @Wikenheiser2013]. Successful integration of the task was further evidenced by the ability of individuals to describe near-optimal strategies after participation, discarding the possibility of our results being driven by noise or bias in the estimation of environmental richness [@Cash-Padgett2020; @Garrett2020; @Dundon2020]. In contrast, our prediction that effort requirements would produce more acceptances only held for the cognitive group, and this tendency for higher acceptances did not switch to passive waiting as predicted for the within-subjects experiment. These hypotheses originally stemmed from prevalent proposals that incentivized work should become more appealing [@Eisenberger1992; @Inzlicht2018], but that people would opt for the easiest available way to achieve desired outcomes [@Hull1943; @Olivola2013]. Instead, our novel use of foraging paradigms, in addition to our inclusion of types of effort that are often studied separately, highlighted task-specific attributes contributing to the attractive or aversive nature of demands. The fact that predicted stable preferences for cognitive effort carried over onto the beginning of the within-subjects experiment further supported a task-specific modulation. However, the unexpected gradual convergence in acceptances in the second experiment emphasized that these features could be contextually molded.

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggested that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discarded the effects of boredom, which would have accentuated the appeal of active tasks over time [**CITE**], or the possibility that our sample was mostly composed of demand seekers [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweis2018], as participants expressed their choices early and rarely quit during the handling time. Risk was also an unlikely factor, as task performance improved over time. Finally, the observed differences in acceptances between effort types could not be explained by reward alone [@Eisenberger1996], as demands were rewarded at the same rate.

  Traditional foraging models, in which a reward is weighed against the earnings foregone by pursuing it [@Charnov1975; @Constantino2015], were also not enough to explain our results. Instead, we computationally found that each type of demand distinctly biased the estimation of this opportunity cost, and that such biases could be gradually recalibrated over time as decision makers gained experienced with additional demands. In this way, waiting and performing physical effort promoted selectivity by adding to the opportunity cost, whereas engaging in cognitive effort boosted an offer's value by reducing it. This raises the question of how these biases arise. We hypothesize that demands altered the perceived duration of time. For example, even though participants knew the nominal handling times, cognitive effort might have become valuable by subjectively compressing them. In our experiments, shorter subjective handling times resulted in higher perceived earning rates, but it is also possible that individuals benefitted from a sense of fast completion. Alterations of subjective time sensitivity are often experienced when we are fully immersed in a task [@Csikszentmihalyi2014], and have been shown to play a role in temporal discounting [@Mckerchar2009]. Such construal of elapsed time could then be adjusted as individuals experienced multiple forms of demands. In fact...Mention that order effects reflect the gradual integration of subjective time across demands in exp 2.

  The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory that impede estimations of the nominal time (Buhi and Gershman), or because participants could count the number of sub-tasks during pursual, which was always less than the number of seconds in the handling time. These hypothesis will have to be experimentally probed in the future.

This would be aking to asking people involved in different tasks to stop after a certain period of time

  (**hold off on this paragraph until modeling is done**) We do not discard the influence of these elements in daily effort evaluations.  make sure you make a big deal out of outcome-independence.

@@

  and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe delay durations, which could hypothetically be recalibrated when tasks are interleaved.
